---
title: "Lecture 8 Inference for Categorical Data"
author: "Rushabh Barbhaya"
date: "September 19, 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Own Your Own

### Question:
The question of atheism was asked by WIN-Gallup International in a similar survey that was conducted in 2005. (We assume here that sample sizes have remained the same.) Table 4 on page 13 of the report summarizes survey results from 2005 and 2012 for 39 countries.

1. Answer the following two questions using the inference function. As always, write out the hypotheses for any tests you conduct and outline the status of the conditions for inference.

    a. Is there convincing evidence that Spain has seen a change in its atheism index between 2005 and 2012? \
    *Hint: Create a new data set for respondents from Spain. Form confidence intervals for the true proportion of athiests in both years, and determine whether they overlap.* \
    b. Is there convincing evidence that the United States has seen a change in its atheism index between 2005 and 2012? \

```{r preRec, results='hide', include=FALSE}
download.file("http://www.openintro.org/stat/data/atheism.RData", destfile = "atheism.RData")
load("atheism.RData")
us12 <- subset(atheism, nationality == "United States" & year == 2012)
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
n <- 1000
p <- seq(0, 1, 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
plot(me ~ p, ylab = "Margin of Error", xlab = "Population Proportion")
p <- 0.1
n <- 1040
p_hats <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
}
```

```{r question1a, warning=FALSE}
spain05 <- subset(atheism, nationality == "Spain" & year == "2005")
spain05$nationality <- as.factor(as.character(spain05$nationality))
table(spain05$nationality, spain05$response)

spain12 <- subset(atheism, nationality == "Spain" & year == "2012")
spain12$nationality <- as.factor(as.character(spain12$nationality))
table(spain12$nationality, spain12$response)

inference(spain05$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")

inference(spain12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```

The assumptions:

$H_0:$ The **number** of atheists in Spain did not change from 2005 to 2012 ($p_{12}=p_6 = 0.1$)
$H_A:$ The **number** changed ($p_{12} \ne 0.1$)\
\
$Solution:$  \
$p_{12} = 0.09$  which is less than 0.1; Therefore, we fail to reject the null hypothesis
\
\
Now, sub-question b.
The assumptions:

$H_0:$ The **number** of atheists in America did not change from 2005 to 2012 ($p_{12}=p_6 = 0.1$)
$H_A:$ The **number** changed ($p_{12} \ne 0.1$)\

```{r question1b, warning=FALSE}
us05 <- subset(atheism, nationality == "United States" & year == "2005")
us05$nationality <- as.factor(as.character(us05$nationality))
table(us05$nationality, us05$response)

us12$nationality <- as.factor(as.character(us12$nationality))
table(us12$nationality, us12$response)

inference(us05$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```
\
$Solution:$  \
$p_{12} = 0.0499$  which is less than 0.1; Therefore, we fail to reject the null hypothesis
\

### Question: \
2. If in fact there has been no change in the atheism index in the countries listed in Table 4, in how many of those countries would you expect to detect a change (at a significance level of 0.05) simply by chance? \
*Hint: Look in the textbook index under Type 1 error.*

$Solution:$ \
Type 1 error says that we rejected a null hypothesis even if the that should not be the case. This can happen purely on chance as well. \
At a significance level of 0.05 with 39 countries we would make a Type 1 error on $39 \times 0.05 = 1.95 \approx 2$ countries. \

### Question:
3. Suppose you're hired by the local government to estimate the proportion of residents that attend a religious service on a weekly basis. According to the guidelines, the estimate must have a margin of error no greater than 1% with 95% confidence. You have no idea what to expect for **\(p\)**. How many people would you have to sample to ensure that you are within the guidelines? \
*Hint: Refer to your plot of the relationship between **\(p\)** and margin of error. Do not use the data set to answer this question.* \
\
$Solution:$ \
If the margin of error is $0.01,$ then at 95% the standard error would be \
$$
SE = \frac{0.01}{1.96} \approx 0.0051
$$
    Assuming, $p = 0.5$, we get
    $$
    SE = \sqrt{\frac{p(1-p)}{n}}
    $$
    $$
    n = \frac{p(1-p)}{SE^2}
    $$
    $$
    n = \frac{0.5\times0.5}{0.01\div1.96}
    $$
    $$
    n = 9604
    $$
    
    We need at the least 9604 samples.